---
title: "Method for updating air screening dispersion factors"
author: "MPCA - Risk evaluation and air modeling, EAO"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Method for updating air screening dispersion factors}
---


## Mission

Default dispersion factors used for screening purposes are updated after the release of a new AERMOD version or MET data set. The model versions^[AERMOD version: 15181] and meteorology^[Met years: 2009-2013; sites MSP & DYT] used for the current dispersion factors are shown on the right. The generated dispersion factors are designed to be protective and overestimate dispersion for real-world operating conditions. The conservative assumptions allow the results to be generalized to a wide set of operation scenarios and provide for quick estimates of air concentrations.


## Updates

The table of default dispersion factors was expanded to include the following options.


- Downwash efffects
- Minimum exit velocity
- Minimum stack temperature
- Minimum stack diameter


_If you have requests or recommendations, please submit them to https://github.com. For general questions, contact [air.modeling@mn.state.us](emailto::air.modeling@mn.state.us)._

# Assumptions

1. __Downwash:__ Worst-case downwash parameters were captured for each scenario by including buildings located 50 meters from the facility in each of the 8 cardinal directions. Three building heights were modeled for each scenario: 50%, 75% and 100% of the stack height.   
1. __Meteorology:__ Each scenario was modeled with met data from `MSP` and `DUL`. 
1. __Stack temperature:__ A minimum stack temperature that is as warm as ambient conditions.  
1. __Exit velocity:__ A minimum exit velocity of 0.2 m/s.

  
_The screening dispersion factors are not considered protective for facilities with exit temperatures below ambient or for sources with near zero exit velocity._

5. __Receptor grid:__  
  > 2 meter spacing for receptors up to 50m from the facility  
  > 5 meter spacing for receptors up to 110m  
  > 10 meter spacing for receptors up to 200m  
  
5. __Flat terrain:__ The model was run with the `flat terrain` option.
5. __Rural:__ The model was run with the default  `rural` option.


# Alernatives

AERSCREEN uses a simple factor method to estimate annual average concentrations - divide the 1-hour maximum by ten. While this method is appropriate for nearby receptors, it results in overly protecive estimates for distances greater than 50 meters.


# Input files
An input file was created for each of the scenario combinations below. `r "672 scenarios were modeled in total."` There was a total of 168 different sources, each modeled using 2 sets of MET data with downwash both and and off.

  - Two met files: Minneapolis (MSP) and Duluth (DYT)   
  - 44 stack heights: [1 to 37 and 40, 45, 50, 60, 70, 80, 90] meters    
  - 7 stack diameters: [0.15, 0.3, 0.6, 1, 1.5, 2.5, 3.5] meters    
  - 6 exit velocities: [0.2, 1, 3, 10, 20, 35] m/s  
  - 4 exit temperatures: [Ambient, 293, 315, 450] kelvin
  - 2: Downwash on/off
 
## Load MET data
```{r eval=F}
library(aermod)

# MSP
MSP <- met_tbl()
MSP$SURFFILE <- "..\\RASSUpdate_Met1\\aq2-115\\MSPMPX5Y_20092013\\MSPMPX5Y.sfc"
MSP$PROFFILE <- "..\\RASSUpdate_Met1\\aq2-115\\MSPMPX5Y_20092013\\MSPMPX5Y.pfl"
MSP$SURFDATA <- "14922 2009 MINNEAPOLIS-ST_PAUL/INTL_ARP"
MSP$UAIRDATA <- "94983 2009"
MSP$PROFBASE <- "256.11 METERS"
MSP$STARTEND <- ""

MSP <- list(MSP = MSP)
                
# DULUTH
DUL <- met_tbl()
DUL$SURFFILE <- "..\\RASSUpdate_Met2\\aq2-104\\DYTINL5Y_20092013\\DYTINL5Y.sfc"
DUL$PROFFILE <- "..\\RASSUpdate_Met2\\aq2-104\\DYTINL5Y_20092013\\DYTINL5Y.pfl"
DUL$SURFDATA <- "4919 2009 Duluth_Sky_ARP"
DUL$UAIRDATA <- "14918 2009 INTERNATIONAL_FALLS/INTL_ARP"
DUL$PROFBASE <- "256.11 METERS"
DUL$STARTEND <- ""

DUL <- list(Duluth = DUL)

```


## Set input parameters
```{r eval=F}

sk_heights <- c(1:37, 40, 45, 50, 60, 70, 80, 90)
  
sk_diameters <- c(0.15, 0.3, 0.6, 1, 1.5, 2.5, 3.5) 
  
sk_velocity <- c(0.2, 1, 3, 10, 20, 35) 

tempsK <- c(0, 293, 315, 450)

tempsF <- list("Ambient", 68, 107, 350)
  
n_scenarios <- 2 * nrow(sk_diameters) * nrow(sk_velocity) * nrow(tempsK) 
  
```

## Run AERMOD for each scenario
```{r eval=F}

for(met_file in list(MSP, DUL)) {

for(height in sk_heights) {

for(diam in sk_diameters) {

for(velocity in sk_velocity) {

for(tempK in tempsK) {

# CONTROL OPTIONS
co <- control_tbl()
co$TITLEONE <- paste(height, "Meter stacks")
co$TITLETWO <- paste0("Met Data: ", names(met_file))
co$MODELOPT <- paste("CONC", "FLAT")
co$AVERTIME <- paste("1", "ANNUAL")

# SOURCE OPTIONS
so <- source_tbl()
so$ELEV     <- strsplit(met_file$PROFBASE, " ")[[1]][1]
so$XCOORD   <- 407502.89
so$YCOORD   <- 5133986.24
so$TYPE     <- "POINT"
so$ID       <- paste(height, grep(diam, sk_diameters), grep(velocity, sk_velocity), grep(tempK, tempsK), sep="_")
so$DESCRSRC <- ""
so$EMISS    <- 1
so$HEIGHT   <- height
so$TEMPK    <- tempK
so$VELOCITY <- velocity
so$DIAMETER <- diam
so$DOWNFILE <- ""
so$GROUPID  <- so$ID
so$GROUPSRC <- so$ID    


# RECEPTORS  
re <- receptor_tbl()
re$RECTFILE <- "..\\Receptors\\circleReceptors.ROU"
    
# METEOROLOGY
me <- metFile[[1]]

# OUTPUT
ou <- out_tbl()
ou$RECTABLE <- "ALLAVE 1ST, 1 1ST"
    
# PROJECT
po <- proj_tbl()
    
# Create master input table
aermod_inp <- cbind(co, po, me)
    

# SAVE INPUT FILE
writeLines(cat(aermod_inp), "aermod.inp")

    
# RUN AERMOD


}}}}}

```

 
# Results

The tables and charts below compare the previous screening dispersion factors to the updated results. 

```{r eval=F}
knitr::kable(
  mtcars[1:6, 1:6], caption = 'A subset of mtcars.'
)
```


`r 'In his later books'`

