---
title: "Updating air screening dispersion factors"
author: "MPCA - Risk evaluation and air modeling, EAO"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Method for updating air screening dispersion factors}
---


## Summary

Default dispersion factors used for screening purposes are updated after the release of a new AERMOD version or MET data set. The most recent model versions^[AERMOD version: 22112] and processed meteorology^[Met years: 2016-2020; sites MSP] were used for the current dispersion factors. The generated dispersion factors are designed to be protective and overestimate dispersion for most real-world operating conditions. The conservative assumptions allow the results to be generalized to a wide set of operation scenarios and provide for quick estimates of air concentrations.


## Updates

The table of default dispersion factors was expanded to include the following options.

- Downwash effects
- Minimum exit velocity
- Minimum stack temperature
- Minimum stack diameter


# Assumptions

1. __Downwash:__ Worst-case downwash parameters were captured for each scenario by including buildings located directly under the source, 20 meters, and 40 meters from the facility in each of the 8 cardinal directions. Three building heights were modeled for each scenario: 75%, 100% and 125% of the stack height.   
1. __Meteorology:__ Each scenario was modeled with meteorology data from `MSP`. 
1. __Stack temperature:__ A minimum stack temperature that is as warm as ambient conditions.  
1. __Exit velocity:__ A minimum exit velocity of 0.2 m/s.

  
_The screening dispersion factors are not considered protective for facilities with exit temperatures below ambient or for sources with near zero exit velocity._

5. __Receptor grid:__  
5. __Flat terrain:__ The model was run with the `flat terrain` option.
5. __Rural:__ The model was run with the default  `rural` option.


# Input files
An input file was created for each of the scenario combinations below.
 
  - 27 stack heights: [3 to 10, and 12,14,16...26, 30,35,40...65, 70,80,90] meters    
  - 7 stack diameters: [0.15, 0.3, 0.6, 1, 1.5, 2, 3] meters    
  - 6 exit velocities: [0.2, 1, 3, 10, 20, 35] m/s  
  - 4 exit temperatures: [Ambient, 293, 315, 450] kelvin
 
 
## Model setup
```{r eval=F}
library(tidyverse)
library(installEPA)
library(aermod)
library(bpip) #remotes::install_github(c("dKvale/receptors", "dKvale/bpip"))

install_epa(c("aermod", "aerscreen", "bpip", "makemet"), 
            dir = "EPA")

```

 
## Set MET data options

```{r eval=F}

# Site specific MET data
## MSP airport is used below
aer_input <- new_aermod()

# Add SFC and PFL files in aermod.exe folder ----
aer_input$surf_file <- "AERMET_KMSPMPX_2016.SFC" 
aer_input$prof_file <- "AERMET_KMSPMPX_2016.PFL" 
aer_input$surf_site_info <- "21112 2016 MINNEAPOLIS-ST_PAUL/INTL_ARP"
aer_input$upper_air_info <- "72649 2016 Chanhassen"
aer_input$base_elev_m <- "0.0 METERS"
aer_input$start_date <- ""
aer_input$end_date <- ""

glimpse(aer_input)
```


## Set receptor options
```{r eval=F}
# Point to receptor grid file
## aer_input$receptor_file <- "receptors.rou"

# Create polar receptor grid ----
aer_input$source_id <- "STK1"

grid_text <- paste0("RE GRIDPOLR POL1 STA \n",
                    "                 ORIG STK1 \n",
                    "                 DIST 10. 50. 100. 300. 500. 1000. 2000. \n",
                    "                 GDIR 36 10. 10. \n",
                    "RE GRIDPOLR POL1 END")

cat(grid_text)

aer_input$receptor_as_text <- grid_text
aer_input$recept_as_text   <- grid_text  # Ensures this works w/ previous vers. of aermod package
```


## Set contstant parameters that remain stable for each run ----
```{r eval=F}
aer_input$x_coord  <- 0
aer_input$y_coord  <- 0
aer_input$type     <- "POINT"
aer_input$desc_src <- ""
aer_input$emit_gs  <- 1
aer_input$elevation_m <- 0
```


## Set input parameter ranges for loop
```{r eval=F}
sk_heights <- c(3:10, seq(12, 26, 2), seq(30, 70, 5), seq(80, 90, 10))

sk_diameters <- c(0.15, 0.3, 0.6, 1, 1.5, 2, 3) 

sk_velocity  <- c(0.5, 1, 3, 10, 20, 35)

tempsK <- c(0, 293, 315, 450)
# tempsF <- list("Ambient", 68, 107, 350)

bld_ht_fractions <- c(0.75, 1, 1.25)


n_scenarios <- length(sk_heights) * length(sk_diameters) * length(sk_velocity) * length(tempsK) * length(bld_ht_fractions) 
```


## Run AERMOD for each scenario ----
```{r eval=F}
for (height in sk_heights) {
  
  # Run BPIP for new stack height
  ## Loop through 3 different building heights
  for (bld_ht_frx in bld_ht_fractions) {
    
    # For stacks > 60 meters, skip to next after first bpip run
    ## Reduces the # of runs with extremely high buildings
    if (height > 60 & bld_ht_frx != bld_ht_fractions[1]) next()  
   
    
    ## Create new bpip table
    builds <- new_bpip()
    
    # Set uniform building params
    builds$source_name   <- "STK1"
    builds$source_height <- height
    builds$bld_height    <- height * bld_ht_frx
    builds$bld_rotation  <- 45
    builds$length_y      <- 15
    builds$width_x       <- 15
    
    glimpse(builds)
  
    
    # Add 16 buildings around the source at
    # 2 distances: 20 and 40 meters
    for (i in 1:16) {
      bld_distance <- ifelse(i < 9, 20, 40)
      
      print(c(i, bld_distance))
      
      builds[i,]                   <- builds[1,]
      builds[i,]$bld_id            <- paste0("BLD_", i)
      builds[i,]$dist_from_source  <- bld_distance 
      builds[i,]$angle_from_source <- (i-1) * 45
    }
    
    # Add a building directly under source
    builds[17, ] <- builds[1, ]
    builds[17, ]$bld_id <- paste0("BLD_", 17)
    builds[17, ]$dist_from_source  <- 0
    builds[17, ]$angle_from_source <- 0
    builds[17, ]$bld_height <- height
    
    glimpse(builds)
    
    # Plot building
    plot_bpip(builds)
    
    # Save bpip input file
    bpi_name <- paste0("input_building_", height, ".bpi")
    
    write_bpip(builds, paste0("EPA/bpip/", bpi_name))
    
    # Run bpip
    run_bpip(input  = bpi_name, 
             output = paste0("downwash_results_", height), 
             exe_folder = "EPA/bpip")
    
    # Read results
    bpip_results <- read_bpip_out(file = paste0("downwash_results_", height, ".out"),
                                  as_text = FALSE)
    
    #cat(bpip_results)
    
    aer_input$downwash_file <- paste0("downwash_results_", height, ".out")
  
    
  # Loop through remaining parameters
  for (diam in sk_diameters) {
    for (velocity in sk_velocity) {
      for (tempK in tempsK) {
        
        # Current scenario
        cat(paste0("Running\n Stack height: ", height,
                     "\n Building height: ", bld_ht_frx,
                     "\n Diameter: ", diam,
                     "\n Velocity: ", velocity,
                     "\n Temp: ", tempK, "\n\n"))
        
        
        # CONTROL OPTIONS
        aer_input$title <- paste(height, "meter stack")
        aer_input$subtitle <- paste0("Met Data: ", aer_input$surf_site_info)
        aer_input$model_opt <- paste("CONC", "FLAT")
        aer_input$aver_time <- paste("1", "24", "ANNUAL")
        
        # SOURCE OPTIONS
        aer_input$height_m <- height
        aer_input$temp_k   <- tempK
        aer_input$velocity_ms <- velocity
        aer_input$diameter_m  <- diam
        aer_input$group_id  <- aer_input$source_id
        aer_input$group_src <- aer_input$source_id    
        
        # DOWNWASH
        ## Set above: Re-ran for each stack height and 3 building heights
        
        # RECEPTORS  
        ## Set globally above as polar grid
          
        # METEOROLOGY
        ## Set globally above
        
        # OUTPUT
        aer_input$rect_table <- "ALLAVE 1ST, 1 1ST, 24 1ST"
          
        # SAVE INPUT FILE
        write_aermod(aer_input, path = "aermod.inp")
          
        # RUN AERMOD
        run_aermod(exe_folder = "EPA",
                   out_file = paste0(height, "_", 
                                     velocity, "_",
                                     diam, "_",
                                     tempK, "_",
                                     "bld", 
                                     gsub("[.]", "", bld_ht_frx), 
                                     ".out"))

}}}}}

```


## Process results ----
```{r eval=F}
# Blank results table
results <- tibble()

distances <- c(seq(10, 99, 1), seq(100, 2000, 100))

avg_times <- c("ANNUAL AVERAGE CONCENTRATION", 
               "1ST HIGHEST  24-HR AVERAGE CONCENTRATION",
               "1ST HIGHEST  1-HR AVERAGE CONCENTRATION")

# Loop through each of the aermod output files
for (i in list.files("./your_output_directory", 
                     pattern = ".out", 
                     full.names = TRUE)) {
  
  # Create table to store result for each distance and avg time
  tmp_results <- tibble(distance = rep(distances, each = length(avg_times)),
                        avg_time = rep(avg_times, length(distances)),
                        disp_factor = as.numeric(NA))
  
  df <- read_aermod_out(i)
  
  df$distance <- sqrt(df$x_coord**2 + df$y_coord**2)
  
  # Find max value for each buffer distance
  for (x in 1:nrow(tmp_results)) {
    
    tmp_row <- tmp_results[x, ]
    
    max_disp_factor <- df %>%
      filter(distance >= tmp_row$distance,
             model_opts == tmp_row$avg_time) %>%
      .$concentration %>%
      max(., na.rm = TRUE)
    
    tmp_results[x, "disp_factor"] <- max_disp_factor
  }
  
  results <- bind_rows(tmp_results, results)
}
```
 
